---
title: "Project proposal"
author: "MDR Intertidal"
output: github_document
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(readxl)

```

## 1. Introduction

Our general research question is "How has the Mount Desert Rock intertidal zone changed over time?". We will be using indicator species such as periwinkles to measure changes in their population. We will also look at changes in seaweed cover throughout the years.

We will be using intertidal data from the Northeastern Coastal Station Alliance, a collection of field stations in the Gulf of Maine. We will specifically be analysing data collected on Mount Desert Rock. It consists of vertical and horizontal transects collected through quadrats in the intertidal zone. For this project we will be using the horizontal transects. This data was collected annually from 2017 to 2023. In the early years, the data was collected by high school students who were trained on the rock. The project was continued by Marina in 2022 and 2023. 

## 2. Data

```{r}
<- read_excel(path = "data/", )
```

Place your data in the /data folder, and add dimensions and codebook to the README in that folder. Then print out the output of glimpse() or skim() of your data frame.

## 3. Data analysis plan

Very preliminary exploratory data analysis, including some summary statistics and visualizations, along with some explanation on how they help you learn more about your data. (You can add to these later as you work on your project.)
The data visualization(s) that you believe will be useful in exploring your question(s). (You can update these later as you work on your project.)


The variables that we will be using to explore our data are tide hight, year, species, sea water temperature, and ascophyllum length. We do not anticipate needing to find any more data to help with this research question.



