---
title: "Project proposal"
author: "MDR Intertidal"
output: github_document
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(readxl)

```

## 1. Introduction

Our general research question is "How has the Mount Desert Rock intertidal zone changed over time?". We will be using indicator species such as periwinkles to measure changes in their population. We will also look at changes in seaweed cover throughout the years.

We will be using intertidal data from the Northeastern Coastal Station Alliance, a collection of field stations in the Gulf of Maine. We will specifically be analysing data collected on Mount Desert Rock. It consists of vertical and horizontal transects collected through quadrats in the intertidal zone. For this project we will be using the horizontal transects. This data was collected annually from 2017 to 2023. In the early years, the data was collected by high school students who were trained on the rock. The project was continued by Marina in 2022 and 2023. 

## 2. Data

```{r load-horizontal-data}

horizontal_MDR_NeCSA_2017  <- read_excel("/cloud/project/data/2017_horizontal_MDR_NeCSA.xlsx")

horizontal_L_MDR_NeCSA_2018 <- read_excel("/cloud/project/data/2018_full_NeCSA.xlsx", sheet = "West HT Low")

horizontal_M_MDR_NeCSA_2018 <- read_excel("/cloud/project/data/2018_full_NeCSA.xlsx", sheet = "West HT Mid")

horizontal_H_MDR_NeCSA_2018 <- read_excel("/cloud/project/data/2018_full_NeCSA.xlsx", sheet = "West HT High")

# horizontal_MDR_NeCSA_2019

# horizontal_MDR_NeCSA_2020


horizontal_MDR_NeCSA_2021  <- read_excel("/cloud/project/data/2021_horizontal_MDR_NeCSA.xlsx")

horizontal_MDR_NeCSA_2022  <- read_excel("/cloud/project/data/2022_horizontal_MDR_NeCSA.xlsx")

horizontal_MDR_NeCSA_2023  <- read_excel("/cloud/project/data/2023_horizontal_MDR_NeCSA.xlsx")

```


```{r - load temperature data}
MDR_temp_20170812_20180811 <- read_excel("/cloud/project/data/MDR_temperature_2017_2021.xlsx", sheet = "20170812-20180811")

MDR_temp_20180830_20180929 <- read_excel("/cloud/project/data/MDR_temperature_2017_2021.xlsx", sheet = "20180830-20180929")

MDR_temp_20180812_20190905 <- read_excel("/cloud/project/data/MDR_temperature_2017_2021.xlsx", sheet = "20180812-20190905")

# MDR_temp_2020

WC_MDR_temp_20210602_20211008 <- read_excel("/cloud/project/data/MDR_temperature_2017_2021.xlsx", sheet = "20210602-20211008WC")
```
```{r}
glimpse(horizontal_MDR_NeCSA_2017)
glimpse(horizontal_L_MDR_NeCSA_2018)
glimpse(horizontal_M_MDR_NeCSA_2018)
glimpse(horizontal_H_MDR_NeCSA_2018)
glimpse(horizontal_MDR_NeCSA_2021)
glimpse(horizontal_MDR_NeCSA_2022)
glimpse(horizontal_MDR_NeCSA_2023)

glimpse(MDR_temp_20170812_20180811)
glimpse(MDR_temp_20180830_20180929)
glimpse(MDR_temp_20180812_20190905)
glimpse(WC_MDR_temp_20210602_20211008)
```

## Loading Data Elaboration

We are awaiting access to the 2019 and 2020 horizontal survey datasets.

There are two sites for temperature sensor attachment on the island - one in the north cove and one in the west cove. We will be using the west cove sensor data as we are exclusively analyzing the west cove location for the horizontal quadrat survey. Only one of the sheets on the datasheet has a specification for which sensor location. We have requested clarification from the manager of the dataset. Additionally, we have requested access to the 2020 portion of the data.



## 3. Data analysis plan

The variables that we will be using to explore our data are tide height, year, species, sea water temperature, and ascophyllum length. We do not anticipate needing to find any more data to help with this research question.

For the preliminary data analysis, we will look at the mean tide height for certain indicator species. We also plan on mutating the data to be able to compare the ratio of crabs to snails over time. In addition to that, we will look at how the percentage of canopy cover affects the presence of snails and crabs across the quadrats. Finally, we will create a bar graph of indicator species abundance faceted by year.


